Omix

### 2019-04-12 omix/omix.Rmd
functionalized most data processing and plotting of omics features{

```{r load_libraries}
library(circlize)
library(ComplexHeatmap)
library(crunch)
library(ggpubr)
library(ggrepel)
library(ggsignif)
library(kableExtra)
library(naniar)
library(openxlsx)
library(reshape2)
library(tidyverse)
library(VennDiagram)

theme_set(theme_light())
```

```{r functions}
unfactorize <- function(df){
  for(i in which(sapply(df, class) == "factor")) {
    df[[i]] = as.character(df[[i]])
  }
  return(df)
}
```

@@

DepMap 19Q1 update fixed almost all of the inconsistencies of the Primary.Disease column, though there were come capitalization inconsistencies.
```{r ccl_info_processing, eval=FALSE}
ccl_info <- read.delim("./data/DepMap/DepMap-2019q1-celllines_v2_20190305.csv", sep = ",", header = TRUE, na.strings = c("", NA))

@@

Mutation calls (coding region, germline filtered):
```{r mutation, eval=FALSE}
maf_raw <- read.delim("./../data_munging/19Q1/depmap_19Q1_mutation_calls_20190305.csv.gz", header = TRUE, sep = ",", row.names = 1)
# Harmonize ID column name
colnames(maf_raw)[colnames(maf_raw) == "DepMap_ID"] <- "Broad_ID"
colnames(maf_raw)[colnames(maf_raw) == "Hugo_Symbol"] <- "Gene"
# Merge CCL metadata
maf_raw <- merge(select(ccl_histologies, Broad_ID, TCGA_classification), maf_raw, by = "Broad_ID", all.y = TRUE)
# Select columns
maf_df <- subset(maf_raw, select = c("Gene", "Broad_ID", "TCGA_classification", "Variant_Classification", "Reference_Allele", "Tumor_Seq_Allele1", "Genome_Change"))
maf_df$Reference_Allele <- ifelse(maf_df$Reference_Allele == "-", 0, nchar(as.character(maf_df$Reference_Allele)))
maf_df$Tumor_Seq_Allele1 <- nchar(as.character(maf_df$Tumor_Seq_Allele1))
maf_df <- maf_df[maf_df$Variant_Classification != "Silent",]
# Add Mutation_Status column
maf_df$Mutation_Status <- 1

@@

DepMap WES CN Data:
```{r copy_number, eval=FALSE}
cn <- read.delim("./../data_munging/19Q1/public_19Q1_gene_cn_20190305.csv.gz", sep = ",", check.names = FALSE, header = TRUE)
# Remove Entrez gene IDs from colnames, format is "Hugo_Symbol (Entrez ID)"
colnames(cn) <- gsub(" .*", "", colnames(cn))
colnames(cn)[1] <- "Broad_ID"
# Merge CCL metadata with TCGA labels
cn <- merge(select(ccl_histologies, Broad_ID, TCGA_classification), cn, by = "Broad_ID", all.y = TRUE)
# Melt
cn_melt <- melt(data = cn, id.vars = c("Broad_ID", "TCGA_classification"), measure.vars = colnames(cn[3:ncol(cn)]), variable.name = "Gene", value.name = "CN")
saveRDS(cn_melt, "./../data_munging/19Q1/cn_melt_19Q1.rds", compress = "xz")
```

19Q1, CCLE RNAseq gene expression data (log2(TPM+1)):
```{r gene_expression, eval=FALSE}
ge <- read.delim("./../../htscreens_giant_files/CCLE_depMap_19Q1_TPM_20190305.csv.gz", header = TRUE, sep = ",", check.names = FALSE)
# Remove Entrez gene IDs from colnames, format is "Hugo_Symbol (Entrez ID)"
colnames(ge) <- gsub(" .*", "", colnames(ge))
colnames(ge)[1] <- "Broad_ID"
# Merge CCL metadata with TCGA labels
ge <- merge(select(ccl_histologies, Broad_ID, TCGA_classification), ge, by = "Broad_ID", all.y = TRUE)
# Melt
ge_melt <- melt(ge, id.vars = c("Broad_ID", "TCGA_classification"), measure.vars = colnames(ge)[3:ncol(ge)], variable.name = "Gene", value.name = "TPM")
saveRDS(ge_melt, "./../data_munging/19Q1/ge_melt_19Q1.rds", compress = "xz")
```

@@

```{r crispr, eval=TRUE}
# Sample info from figshare
crispr_meta <- read.delim("./data/DepMap/sample_info_20190306.csv", sep = ",", header = TRUE, na.strings = c("", NA))
colnames(crispr_meta)[colnames(crispr_meta) == "CCLE_name"] <- "CCLE_Name"

@@

```{r pozdeyev_processing, eval=FALSE}
ccle_poz <- read.delim("./data/pozdeyev2016/CCLE_drug_response_metrics.csv.gz", sep = ",")
# ccle_test <- merge(select(ccle_poz, Drug_name, Cell_line, AUC_IC50), select(filter(auc, Database == "CCLE"), Drug_name, Cell_line, AUC_IC50), by = c("Drug_name", "Cell_line"), suffixes = c("_unmatched", "_matched_all_pairs"), all = TRUE)

@@

```{r exploratory_plots, eval=FALSE}
# Data subsetting for easier tracking throughout plots
pgx_ccls <- filter(auc_all, WasDrugScreened == 1)
crispr_ccls <- filter(auc_all, WasCRISPRScreened == 1)
unscreened_ccls <- filter(auc_all, WasCRISPRScreened == 0 & WasDrugScreened == 0)
ge_ccls <- filter(auc_all, HasGeneExpressionData == 1)
cn_ccls <- filter(auc_all, HasCopyNumberData == 1)
mut_ccls <- filter(auc_all, HasMutationData == 1)
ccle_ccls <- filter(auc_all, InCCLE == 1)
ctrp_ccls <- filter(auc_all, InCTRP == 1)
gdsc_ccls <- filter(auc_all, InGDSC == 1)

@@

@@

Data wrangling and processing:
```{r ge_heatmap_data, eval=FALSE, warning=FALSE}
# DepMap gene expression
ge_depmap <- read.delim("./../../htscreens_giant_files/omix/GE/GE_DepMap_filter.csv.gz", sep = ",", header = TRUE, check.names = FALSE)
colnames(ge_depmap)[colnames(ge_depmap) == "cancer"] <- "TCGA_classification"
# Mean center scale
ge_depmap[, 4:ncol(ge_depmap)] <- apply(ge_depmap[, 4:ncol(ge_depmap)], MARGIN = 2, FUN = scale, center = TRUE, scale = TRUE)
# Melt
ge_depmap_melt <- melt(ge_depmap, id.vars = c("sample_ID", "TCGA_classification"), measure.vars = colnames(ge_depmap)[4:length(colnames(ge_depmap))], variable.name = "Gene", value.name = "TPM_scaled")
# Average scaled TPM per gene per TCGA classification
ge_depmap_avg <- ge_depmap_melt %>% group_by(TCGA_classification, Gene) %>% summarise(Mean_TPM = mean(TPM_scaled))
# Make wide
ge_depmap_wide <- dcast(data = ge_depmap_avg, formula = Gene ~ TCGA_classification, fun.aggregate = mean)
rownames(ge_depmap_wide) <- ge_depmap_wide$Gene
ge_depmap_wide$Gene <- NULL
colnames(ge_depmap_wide) <- paste(colnames(ge_depmap_wide), "DepMap", sep = "_")

# TCGA gene expression
ge_tcga <- read.delim("./../../htscreens_giant_files/omix/GE/GE_Xena_filter.csv.gz", sep = ",", header = TRUE, check.names = FALSE)
colnames(ge_tcga)[colnames(ge_tcga) == "cancer"] <- "TCGA_classification"
# Mean center scale
ge_tcga[, 4:ncol(ge_tcga)] <- apply(ge_tcga[, 4:ncol(ge_tcga)], MARGIN = 2, FUN = scale, center = TRUE, scale = TRUE)
# Melt
ge_tcga_melt <- melt(ge_tcga, id.vars = c("sample_ID", "TCGA_classification"), measure.vars = colnames(ge_tcga)[4:length(colnames(ge_tcga))], variable.name = "Gene", value.name = "TPM_scaled")
# Average scaled TPM per gene per TCGA classification
ge_tcga_avg <- ge_tcga_melt %>% group_by(TCGA_classification, Gene) %>% summarise(Mean_TPM = mean(TPM_scaled))
# Make wide
ge_tcga_wide <- dcast(data = ge_tcga_avg, formula = Gene ~ TCGA_classification, fun.aggregate = mean)
rownames(ge_tcga_wide) <- ge_tcga_wide$Gene
ge_tcga_wide$Gene <- NULL
colnames(ge_tcga_wide) <- paste(colnames(ge_tcga_wide), "TCGA", sep = "_")

# Correlate all against all
ge_wide <- bind_cols(ge_depmap_wide, ge_tcga_wide)
ge_corr <- ge_wide %>% as.matrix %>% cor(method = "spearman") %>% as.data.frame %>% rownames_to_column(var = "var1") %>% gather(var2, value, -var1)
# Filter self-comparisons and eliminate duplicate correlations
ge_corr <- filter(ge_corr, !(grepl("DepMap", var1) & grepl("DepMap", var2)) & !(grepl("TCGA", var1) & grepl("TCGA", var2)) & grepl("TCGA", var2))
# Data management
colnames(ge_corr) <- c("DepMap", "TCGA", "r")
ge_corr <- data.frame(apply(ge_corr, 2, function(x) gsub("_TCGA|_DepMap","", x)))
ge_corr$r <- as.numeric(as.character(ge_corr$r))
# Set up coprrelation dataframe for plotting
ge_corr_wide <- dcast(ge_corr, DepMap ~ TCGA)
rownames(ge_corr_wide) <- ge_corr_wide$DepMap
ge_corr_wide$DepMap <- NULL
```

@@

Make heatmaps:
```{r ge_heatmap, eval=FALSE, fig.height=4, fig.width=3.75}
col_fun <- colorRamp2(c(-0.5, 0, 0.5, 1), c("black", "red3", "yellow", "white"))
ge_ht4order <- Heatmap(ge_corr_wide, cluster_rows = FALSE)

ge_ht_legend_list <- list(legend_direction = "horizontal", title_position = "lefttop", legend_width = unit(1.5, "in"), border = "black")

ge_ht <- Heatmap(ge_corr_wide, 
                 name = "Expression correlation",
                 col = col_fun, 
                 rect_gp = gpar(col = "black"),
                 heatmap_legend_param = ge_ht_legend_list,
                 row_order = rev(column_order(ge_ht4order)), 
                 show_column_dend = FALSE, 
                 cluster_rows = FALSE, 
                 row_names_side = "left",
                 row_title = "DepMap cancer cell lines",
                 column_title = "TCGA tumor samples", 
                 column_title_side = "bottom")
png(filename = "./figures/ge_depmap_tcga_heatmap_spearman.png", width = 7.5, height = 8, units = "in", res = 500)
draw(ge_ht, heatmap_legend_side = "top")
dev.off()
```

@@

Data wrangling and processing:
```{r cn_heatmap_data, eval=FALSE, warning=FALSE}
# TCGA copy number
cn_tcga <- read.delim("./data/CN/CN_GDC_filter.csv.gz", sep = ",", header = TRUE, check.names = FALSE)
colnames(cn_tcga)[colnames(cn_tcga) == "cancer"] <- "TCGA_classification"
# Melt
cn_tcga_melt <- melt(cn_tcga, id.vars = c("sample_ID", "TCGA_classification"), measure.vars = colnames(cn_tcga)[4:length(colnames(cn_tcga))], variable.name = "Gene", value.name = "CN")
# Average scaled TPM per gene per TCGA classification
cn_tcga_avg <- cn_tcga_melt %>% group_by(TCGA_classification, Gene) %>% summarise(Mean_CN = mean(CN))
# Make wide
cn_tcga_wide <- dcast(data = cn_tcga_avg, formula = Gene ~ TCGA_classification, fun.aggregate = mean)
rownames(cn_tcga_wide) <- cn_tcga_wide$Gene
cn_tcga_wide$Gene <- NULL
colnames(cn_tcga_wide) <- paste(colnames(cn_tcga_wide), "TCGA", sep = "_")

# DepMap copy number (18Q4)
cn_depmap <- read.delim("./data/CN/CN_DepMap_filter.csv.gz", sep = ",", header = TRUE, check.names = FALSE)
colnames(cn_depmap)[colnames(cn_depmap) == "cancer"] <- "TCGA_classification"
# Melt
cn_depmap_melt <- melt(cn_depmap, id.vars = c("sample_ID", "TCGA_classification"), measure.vars = colnames(cn_depmap)[4:length(colnames(cn_depmap))], variable.name = "Gene", value.name = "CN")
# Average CN levels per gene per TCGA classification
cn_depmap_avg <- cn_depmap_melt %>% group_by(TCGA_classification, Gene) %>% summarise(Mean_CN = mean(CN))
# Make wide
cn_depmap_wide <- dcast(data = cn_depmap_avg, formula = Gene ~ TCGA_classification, fun.aggregate = mean)
rownames(cn_depmap_wide) <- cn_depmap_wide$Gene
cn_depmap_wide$Gene <- NULL
colnames(cn_depmap_wide) <- paste(colnames(cn_depmap_wide), "DepMap", sep = "_")

>>>>

```{r load_libraries}
library(circlize)
library(ComplexHeatmap)
library(crunch)
library(ggpubr)
library(ggrepel)
library(ggsignif)
library(kableExtra)
library(lemon)
library(naniar)
library(openxlsx)
library(reshape2)
library(tidyverse)
library(VennDiagram)

theme_set(theme_light())
# install.packages("lemon")
```

```{r functions}
unfactorize <- function(df){
  for(i in which(sapply(df, class) == "factor")) {
    df[[i]] = as.character(df[[i]])
  }
  return(df)
}

DepMapTCGAMakeWideDF <- function(melted_df, output_type, data_source, exclusions) {
  # Input is a 4-column melted dataframe of omic features
  # Rename columns: sample_ID, TCGA_classification, Gene, Value
  colnames(df) <- c("sample_ID", "TCGA_classification", "Gene", "Value")
  if(tolower(output_type) == "average" | tolower(output_type) == "avg") {
    # Average feature values per gene per TCGA classification
    df_avg <- df %>%
              group_by(TCGA_classification, Gene) %>%
              summarise(Mean_Value = mean(Value))
    # Make wide
    df_wide <- dcast(data = df_avg,
                     formula = Gene ~ TCGA_classification,
                     value.var = "Mean_Value",
                     fun.aggregate = mean)
    # Remove genes if argument exists
    if(!missing(exclusions)) { df_wide <- df_wide[df_wide$Gene != exclusions, ] }
    rownames(df_wide) <- df_wide$Gene
    df_wide$Gene <- NULL
    colnames(df_wide) <- paste(colnames(df_wide), data_source, sep = "_")
  }  
  if(tolower(output_type) == "proportion" | tolower(output_type) == "prop") {
    # Calc proportion of 1s over 1s + 0s per gene per TCGA classification
    df_prop <- df %>% 
               group_by(TCGA_classification, Gene) %>%
               summarise(Proportion = sum(Value) / n())
    # Make wide
    df_wide <- dcast(data = df_prop,
                     formula = Gene ~ TCGA_classification,
                     value.var = "Proportion",
                     fun.aggregate = mean)
    # Remove genes if argument exists
    if(!missing(exclusions)) { df_wide <- df_wide[df_wide$Gene != exclusions, ] }
    rownames(df_wide) <- df_wide$Gene
    df_wide$Gene <- NULL
    colnames(df_wide) <- paste(colnames(df_wide), data_source, sep = "_")
  }
  return(df_wide)
}

DepMapTCGACorrMatrix <- function(DepMap_wide_df, TCGA_wide_df, corr_method) {
  # Input is the output from DepMapTCGAMakeWideDF function for DepMap and TCGA
  wide <- bind_cols(DepMap_wide_df, TCGA_wide_df)
  # Correlate all against all
  corr <- wide %>% as.matrix %>% cor(method = corr_method) %>% as.data.frame %>% rownames_to_column(var = "var1") %>% gather(var2, value, -var1)
  # Filter self-comparisons and eliminate duplicate correlations
  corr <- filter(corr, !(grepl("DepMap", var1) & grepl("DepMap", var2)) & !(grepl("TCGA", var1) & grepl("TCGA", var2)) & grepl("TCGA", var2))
  # Data management
  colnames(corr) <- c("DepMap", "TCGA", "r")
  corr <- data.frame(apply(corr, 2, function(x) gsub("_TCGA|_DepMap","", x)))
  corr$r <- as.numeric(as.character(corr$r))
  # Set up coprrelation dataframe for plotting
  corr_wide <- dcast(corr, DepMap ~ TCGA, value.var = "r")
  rownames(corr_wide) <- corr_wide$DepMap
  corr_wide$DepMap <- NULL
  return(corr_wide)
}

DepMapTCGACorrHeatmap <- function(corr_wide_df, out_filepath, legend_title, color_breaks) {
  # Libraries
  require(ComplexHeatmap)
  require(circlize)
  # Define color scale
  col_fun <- colorRamp2(color_breaks, c("black", "red3", "yellow", "white"))
  # Legend properties list
  ht_legend_list <- list(legend_direction = "horizontal", title_position = "lefttop", legend_width = unit(1.5, "in"), border = "black")
  # Get column order to have symmetrically ordered axes
  ht4order <- Heatmap(corr_wide_df, cluster_rows = FALSE)
  # Make heatmap
  ht <- Heatmap(corr_wide_df, 
                name = legend_title,
                col = col_fun, 
                rect_gp = gpar(col = "black"),
                heatmap_legend_param = ht_legend_list,
                row_order = rev(column_order(ht4order)), 
                show_column_dend = FALSE, 
                cluster_rows = FALSE, 
                row_names_side = "left",
                row_title = "DepMap cancer cell lines",
                column_title = "TCGA tumor samples", 
                column_title_side = "bottom")
  # Save to file
  png(filename = out_filepath, width = 7.5, height = 8, units = "in", res = 500)
  draw(ht, heatmap_legend_side = "top")
  dev.off()
  # Return heatmap object for in-line rendering
  return(ht)
}

DepMapTCGAPropsScatterplots <- function(DepMap_melted_df, TCGA_melted_df, out_filepath) {
  # Input is the 4-column melted dataframe of omic features from DepMap and TCGA
  # Rename columns: sample_ID, TCGA_classification, Gene, Value
  colnames(DepMap_melted_df) <- c("sample_ID", "TCGA_classification", "Gene", "Value")
  colnames(TCGA_melted_df) <- c("sample_ID", "TCGA_classification", "Gene", "Value")
  # Calculate proportions
  DepMap_prop <- DepMap_melted_df %>% group_by(TCGA_classification, Gene) %>% summarise(Proportion = sum(Value) / n())
  TCGA_prop <- TCGA_melted_df %>% group_by(TCGA_classification, Gene) %>% summarise(Proportion = sum(Value) / n())
  # Merge proportions from DepMap and TCGA
  props <- merge(DepMap_prop, TCGA_prop, by = c("TCGA_classification", "Gene"), all = TRUE, suffixes = c("_DepMap","_TCGA"))
  # Plot faceted scatterplots
  props_scatters <- ggplot(data = props, aes(x = Proportion_TCGA, y = Proportion_DepMap)) +
    facet_wrap(~ TCGA_classification, nrow = 4) +
    geom_point(alpha = 0.3, color = "darkcyan") +
    scale_y_continuous(breaks = seq(0, 1, by = 0.2), labels = seq(0, 1, by = 0.2)) +
    scale_x_continuous(breaks = seq(0, 1, by = 0.2), labels = seq(0, 1, by = 0.2)) +
    coord_cartesian(ylim = c(0, 1), xlim = c(0, 1)) +
    geom_smooth(method = "lm", se = FALSE, lwd = 0.5, color = "darkgray") +
    stat_cor(method = "pearson", label.x = 0.85, label.y = 0.15, label.sep = "\n") +
    labs(x = "TCGA frequency", y = "DepMap frequency", caption = paste0(formatC(x = length(unique(props$Gene)), format = "d", big.mark = ","), " genes compared."))
  # Save to file
  ggsave(out_filepath, props_scatters, device = "png", dpi = 450, width = 12, height = 12, units = "in")
  # Return for in-line rendering
  return(props_scatters)
}
```

@@

DepMap 19Q1 update fixed almost all of the inconsistencies of the Primary.Disease column, though there were come capitalization inconsistencies.
```{r ccl_info_processing, eval = FALSE}
ccl_info <- read.delim("./data/DepMap/DepMap-2019q1-celllines_v2_20190305.csv", sep = ",", header = TRUE, na.strings = c("", NA))

@@

Mutation calls (coding region, germline filtered):
```{r mutation, eval = FALSE}
maf_raw <- read.delim("./../data_munging/19Q1/depmap_19Q1_mutation_calls_20190305.csv.gz", header = TRUE, sep = ",", row.names = 1)
# Harmonize ID column name
colnames(maf_raw)[colnames(maf_raw) == "DepMap_ID"] <- "Broad_ID"
colnames(maf_raw)[colnames(maf_raw) == "Hugo_Symbol"] <- "Gene"
# Merge CCL metadata
maf_raw <- merge(select(ccl_histologies, Broad_ID, TCGA_classification), maf_raw, by = "Broad_ID", all.y = TRUE)
# Select columns
maf_df <- subset(maf_raw, select = c("Gene", "Broad_ID", "TCGA_classification", "Variant_Classification", "Reference_Allele", "Tumor_Seq_Allele1", "Genome_Change"))
maf_df$Reference_Allele <- ifelse(maf_df$Reference_Allele == "-", 0, nchar(as.character(maf_df$Reference_Allele)))
maf_df$Tumor_Seq_Allele1 <- nchar(as.character(maf_df$Tumor_Seq_Allele1))
maf_df <- maf_df[maf_df$Variant_Classification != "Silent",]
# Add Mutation_Status column
maf_df$Mutation_Status <- 1

@@

DepMap WES CN Data:
```{r copy_number, eval = FALSE}
cn <- read.delim("./../data_munging/19Q1/public_19Q1_gene_cn_20190305.csv.gz", sep = ",", check.names = FALSE, header = TRUE)
# Remove Entrez gene IDs from colnames, format is "Hugo_Symbol (Entrez ID)"
colnames(cn) <- gsub(" .*", "", colnames(cn))
colnames(cn)[1] <- "Broad_ID"
# Merge CCL metadata with TCGA labels
cn <- merge(select(ccl_histologies, Broad_ID, TCGA_classification), cn, by = "Broad_ID", all.y = TRUE)
# Melt
cn_melt <- melt(data = cn, id.vars = c("Broad_ID", "TCGA_classification"), measure.vars = colnames(cn[3:ncol(cn)]), variable.name = "Gene", value.name = "CN")
saveRDS(cn_melt, "./../data_munging/19Q1/cn_melt_19Q1.rds", compress = "xz")
```

19Q1, CCLE RNAseq gene expression data (log2(TPM+1)):
```{r gene_expression, eval = FALSE}
ge <- read.delim("./../../htscreens_giant_files/CCLE_depMap_19Q1_TPM_20190305.csv.gz", header = TRUE, sep = ",", check.names = FALSE)
# Remove Entrez gene IDs from colnames, format is "Hugo_Symbol (Entrez ID)"
colnames(ge) <- gsub(" .*", "", colnames(ge))
colnames(ge)[1] <- "Broad_ID"
# Merge CCL metadata with TCGA labels
ge <- merge(select(ccl_histologies, Broad_ID, TCGA_classification), ge, by = "Broad_ID", all.y = TRUE)
# Melt
ge_melt <- melt(ge, id.vars = c("Broad_ID", "TCGA_classification"), measure.vars = colnames(ge)[3:ncol(ge)], variable.name = "Gene", value.name = "TPM")
saveRDS(ge_melt, "./../data_munging/19Q1/ge_melt_19Q1.rds", compress = "xz")
```

@@

```{r crispr}
# Sample info from figshare
crispr_meta <- read.delim("./data/DepMap/sample_info_20190306.csv", sep = ",", header = TRUE, na.strings = c("", NA))
colnames(crispr_meta)[colnames(crispr_meta) == "CCLE_name"] <- "CCLE_Name"

@@

```{r pozdeyev_processing, eval = FALSE}
ccle_poz <- read.delim("./data/pozdeyev2016/CCLE_drug_response_metrics.csv.gz", sep = ",")
# ccle_test <- merge(select(ccle_poz, Drug_name, Cell_line, AUC_IC50), select(filter(auc, Database == "CCLE"), Drug_name, Cell_line, AUC_IC50), by = c("Drug_name", "Cell_line"), suffixes = c("_unmatched", "_matched_all_pairs"), all = TRUE)

@@

```{r exploratory_plots, eval = FALSE}
# Data subsetting for easier tracking throughout plots
pgx_ccls <- filter(auc_all, WasDrugScreened == 1)
crispr_ccls <- filter(auc_all, WasCRISPRScreened == 1)
unscreened_ccls <- filter(auc_all, WasCRISPRScreened == 0 & WasDrugScreened == 0)
ge_ccls <- filter(auc_all, HasGeneExpressionData == 1)
cn_ccls <- filter(auc_all, HasCopyNumberData == 1)
mut_ccls <- filter(auc_all, HasMutationData == 1)
ccle_ccls <- filter(auc_all, InCCLE == 1)
ctrp_ccls <- filter(auc_all, InCTRP == 1)
gdsc_ccls <- filter(auc_all, InGDSC == 1)

@@

```{r}
tcga_classifications_16 <- c("BLCA", "BRCA", "COADREAD", "ESCA", "GBMLGG", "HNSC", "KIRC", "LIHC", "LUAD", "LUSC", "OV", "PAAD", "SARC", "SKCM", "STAD", "UCEC")
```

@@

Data wrangling and processing:
```{r ge_heatmap_data, eval=FALSE, warning=FALSE}
# DepMap
# TPM
ge_depmap <- read.delim("./../../htscreens_giant_files/omix/GE/GE_DepMap_filter.csv.gz", sep = ",", header = TRUE, check.names = FALSE)
colnames(ge_depmap)[colnames(ge_depmap) == "cancer"] <- "TCGA_classification"
# Mean-center scale
ge_depmap[, 4:ncol(ge_depmap)] <- apply(ge_depmap[, 4:ncol(ge_depmap)], MARGIN = 2, FUN = scale, center = TRUE, scale = TRUE)
# Melt
ge_depmap_melt <- melt(ge_depmap, id.vars = c("sample_ID", "TCGA_classification"), measure.vars = colnames(ge_depmap)[4:length(colnames(ge_depmap))], variable.name = "Gene", value.name = "TPM_scaled")
# Make wide
ge_depmap_wide <- DepMapTCGAMakeWideDF(df = ge_depmap_melt, output_type = "average", data_source = "DepMap")

# TCGA gene expression
ge_tcga <- read.delim("./../../htscreens_giant_files/omix/GE/GE_Xena_filter.csv.gz", sep = ",", header = TRUE, check.names = FALSE)
colnames(ge_tcga)[colnames(ge_tcga) == "cancer"] <- "TCGA_classification"
# Mean center scale
ge_tcga[, 4:ncol(ge_tcga)] <- apply(ge_tcga[, 4:ncol(ge_tcga)], MARGIN = 2, FUN = scale, center = TRUE, scale = TRUE)
# Melt
ge_tcga_melt <- melt(ge_tcga, id.vars = c("sample_ID", "TCGA_classification"), measure.vars = colnames(ge_tcga)[4:length(colnames(ge_tcga))], variable.name = "Gene", value.name = "TPM_scaled")
# Make wide
ge_tcga_wide <- DepMapTCGAMakeWideDF(df = ge_tcga_melt, output_type = "average", data_source = "TCGA")

# Correlate all against all
ge_corr_wide_pearson <- DepMapTCGACorrMatrix(DepMap_wide_df = ge_depmap_wide, TCGA_wide_df = ge_tcga_wide, corr_method = "pearson")
ge_corr_wide_spearman <- DepMapTCGACorrMatrix(DepMap_wide_df = ge_depmap_wide, TCGA_wide_df = ge_tcga_wide, corr_method = "spearman")
```

@@

Make heatmaps:
```{r ge_heatmap, eval=FALSE, fig.height=4, fig.width=3.75}
ge_ht_p <- DepMapTCGACorrHeatmap(corr_wide_df = ge_corr_wide_pearson, out_filepath = "./figures/ge_depmap_tcga_heatmap_pearson.png", legend_title = "Expression Pearson correlation", color_breaks = c(-0.5, 0, 0.5, 1))
# draw(ge_ht_p, heatmap_legend_side = "top")

ge_ht_s <- DepMapTCGACorrHeatmap(corr_wide_df = ge_corr_wide_spearman, out_filepath = "./figures/ge_depmap_tcga_heatmap_spearman.png", legend_title = "Expression Spearman correlation", color_breaks = c(-0.5, 0, 0.5, 1))
# draw(ge_ht_s, heatmap_legend_side = "top")
```

Data wrangling and processing:
```{r mut_heatmap_data, eval=FALSE}
# DepMap
mut_depmap <- read.delim("./data/mutation/mutation_DepMap_filter.csv.gz", sep = ",", header = TRUE, check.names = FALSE)
mut_depmap <- unfactorize(mut_depmap)
mut_depmap[, 4:ncol(mut_depmap)] <- apply(mut_depmap[, 4:ncol(mut_depmap)], 2, function(x) ifelse(x > 1, 1, x))
colnames(mut_depmap)[colnames(mut_depmap) == "cancer"] <- "TCGA_classification"
# Melt
mut_depmap_melt <- melt(mut_depmap, id.vars = c("sample_ID", "TCGA_classification"), measure.vars = colnames(mut_depmap)[4:length(colnames(mut_depmap))], variable.name = "Gene", value.name = "Mutation_Status")
# Make wide
mut_depmap_wide <- DepMapTCGAMakeWideDF(df = mut_depmap_melt, output_type = "proportion", data_source = "DepMap", exclusions = c("TP53"))

# TCGA
mut_tcga <- read.delim("./data/mutation/mutation_Xena_filter.csv.gz", sep = ",", header = TRUE, check.names = FALSE)
mut_tcga <- unfactorize(mut_tcga)
mut_tcga[, 4:ncol(mut_tcga)] <- apply(mut_tcga[, 4:ncol(mut_tcga)], 2, function(x) ifelse(x > 1, 1, x))
colnames(mut_tcga)[colnames(mut_tcga) == "cancer"] <- "TCGA_classification"
# Melt
mut_tcga_melt <- melt(mut_tcga, id.vars = c("sample_ID", "TCGA_classification"), measure.vars = colnames(mut_tcga)[4:length(colnames(mut_tcga))], variable.name = "Gene", value.name = "Mutation_Status")
# Make wide
mut_tcga_wide <- DepMapTCGAMakeWideDF(df = mut_tcga_melt, output_type = "proportion", data_source = "TCGA", exclusions = c("TP53"))

# Correlate all against all
mut_corr_wide_pearson <- DepMapTCGACorrMatrix(DepMap_wide_df = mut_depmap_wide, TCGA_wide_df = mut_tcga_wide, corr_method = "pearson")
mut_corr_wide_spearman <- DepMapTCGACorrMatrix(DepMap_wide_df = mut_depmap_wide, TCGA_wide_df = mut_tcga_wide, corr_method = "spearman")
```

Pseudo-figure caption: This is comparison of binary gene-level mutation frequency profiles between DepMap cancer cell lines and TCGA tumor samples, both classified by TCGA project. In each dataset and for each TCGA classification, the proportion of samples containing at least one mutation was calculated for all genes shared between the two datasets, and the proportions were correlated against each other. Pearson/Spearman correlations (r) between TCGA and DepMap mutation feature sets are shown in the heatmap.

For plots labeled as "5000", the 5,000 most variable genes analyzed in the gene expression heatmap were selected for the mutation analyses. Alternatively, for the plots labeled as "252", the genes in the TCGA mutation dataset were analyzed if they were mutated in >=4% annotated samples and had a mutation in at least one cancer cell line from DepMap. TP53 was removed from the heatmap, as was done in Barretina et al. (2012). For the scatterplots, TP53 was retained.

Pseudo-caption for scatterplots: This is a comparison of proportions of mutated samples between cancer cell lines and TCGA samples in our 16 TCGA classification categories. Each point corresponds to a single gene in the associated cancer type.

Make plots:
```{r mut_plots, eval=FALSE, fig.height=4, fig.width=3.75}
mut_ht_p <- DepMapTCGACorrHeatmap(corr_wide_df = mut_corr_wide_pearson, out_filepath = "./figures/mut_depmap_tcga_heatmap_pearson_252.png", legend_title = "Mutation Pearson correlation", color_breaks = c(0, 0.3, 0.6, 0.9))
# draw(mut_ht_p, heatmap_legend_side = "top")

mut_ht_s <- DepMapTCGACorrHeatmap(corr_wide_df = mut_corr_wide_spearman, out_filepath = "./figures/mut_depmap_tcga_heatmap_spearman_252.png", legend_title = "Mutation Spearman correlation", color_breaks = c(0, 0.3, 0.5, 0.7))
# draw(mut_ht_s, heatmap_legend_side = "top")

mut_props_scatters <- DepMapTCGAPropsScatterplots(DepMap_melted_df = mut_depmap_melt, TCGA_melted_df = mut_tcga_melt, out_filepath = "./figures/mut_depmap_tcga_scatter_pearson_253.png")
```

@@

### Continuous

Continuous CN estimate data wrangling and processing:
```{r cn_heatmap_data_continuous, eval=FALSE, warning=FALSE}
# DepMap copy number
cn_depmap <- read.delim("./../../htscreens_giant_files/omix/CN/CN_cont_DepMap_filter.csv.gz", sep = ",", header = TRUE, check.names = FALSE)
colnames(cn_depmap)[colnames(cn_depmap) == "cancer"] <- "TCGA_classification"
# Mean center scale
cn_depmap[, 4:ncol(cn_depmap)] <- apply(cn_depmap[, 4:ncol(cn_depmap)], MARGIN = 2, FUN = scale, center = TRUE, scale = TRUE)
# Melt
cn_depmap_melt <- melt(cn_depmap, id.vars = c("sample_ID", "TCGA_classification"), measure.vars = colnames(cn_depmap)[4:length(colnames(cn_depmap))], variable.name = "Gene", value.name = "CN_scaled")
cn_depmap_melt <- filter(cn_depmap_melt, TCGA_classification %in% tcga_classifications_16)
# Make wide
cn_depmap_wide <- DepMapTCGAMakeWideDF(df = cn_depmap_melt, output_type = "average", data_source = "DepMap")

# TCGA cop number
cn_tcga <- read.delim("./../../htscreens_giant_files/omix/CN/CN_cont_Xena_filter.csv.gz", sep = ",", header = TRUE, check.names = FALSE)
colnames(cn_tcga)[colnames(cn_tcga) == "cancer"] <- "TCGA_classification"
# Mean center scale
cn_tcga[, 4:ncol(cn_tcga)] <- apply(cn_tcga[, 4:ncol(cn_tcga)], MARGIN = 2, FUN = scale, center = TRUE, scale = TRUE)
# Melt
cn_tcga_melt <- melt(cn_tcga, id.vars = c("sample_ID", "TCGA_classification"), measure.vars = colnames(cn_tcga)[4:length(colnames(cn_tcga))], variable.name = "Gene", value.name = "CN_scaled")
cn_tcga_melt <- filter(cn_tcga_melt, TCGA_classification %in% tcga_classifications_16)
# Make wide
cn_tcga_wide <- DepMapTCGAMakeWideDF(df = cn_tcga_melt, output_type = "proportion", data_source = "TCGA")

# Correlate all against all
cn_corr_wide_pearson <- DepMapTCGACorrMatrix(DepMap_wide_df = cn_depmap_wide, TCGA_wide_df = cn_tcga_wide, corr_method = "pearson")
cn_corr_wide_spearman <- DepMapTCGACorrMatrix(DepMap_wide_df = cn_depmap_wide, TCGA_wide_df = cn_tcga_wide, corr_method = "spearman")
```

### -1, 0, +1
